---
title: "Results from CU"
author: "Michelle R. Jackson"
date: "5/1/2018"
output: html_document
---
#################################
#Name: Christine Urbanowicz
#Date: April 9, 2018
#R version: 3.3.1
#Purpose: Code cleans up data by changing prunse cot to fraxinus,
#calculates diversity of native plant species, 
#calculates abundance of fxnl groups, 
#runs PERMANOVA, and creates NMDS plot.
#################################
```{r}
library(vegan) #for adonis and nmds
library(dplyr) #for summarizing
library(reshape2) #for melting and casting
```
##### Open data#####
#earthworm data from Michelle, March 2018
```{r}
wormData<-read.csv("earthwormData.csv")
```
#functional group look-up table, derived from master sheet from Erin, February 2018
```{r}
fxnLUT<-read.csv("fxnlGroupByCode.csv")
```
#plants from Erin - christine cleaned up by removing canopy trees
```{r}
allPlants<-read.csv("allCommunity_canopyRemoved_Feb2018.csv")
names(allPlants)
```
##### Clean data ######
#subset by sites in earthworm dataset
```{r}
allPlants_just2Sites<-allPlants[allPlants$site=="HF"|
                                            allPlants$site=="MC",]
```
#subset by 2017 and remove sprayed treatment
```{r}
allPlants_just2Sites2017<-allPlants_just2Sites[allPlants_just2Sites$year=="2017"&allPlants_just2Sites$treatment!="s",]
```
#get rid of plants without any presence in the 2 sites in 2017
```{r}
community<-allPlants_just2Sites2017[,-c(1:4)]
community_ew<-community[,(colSums(community)!=0)]
```
#take out unknown seedling
```{r}
community_ew$unk_seedling <- NULL
```
#change prunse cotelydon to fraxam!!!!!!!!!!!!! prunse cotyeldon was actually white ash (Erin identified April 2018)
```{r}
community_ew$fraxam<-community_ew$fraxam+community_ew$prunse_cot
```
#get rid of prunse_cot
```{r}
community_ew$prunse_cot <- NULL
```
#double check there are 36 taxa (took away unknown seedling and prunse)
```{r}
ncol(community_ew) ==35 
```
###### Calculate native total stem and shannon diversity #######
#get rid of allipe columns and only other invasive, euonal (burning bursh)
```{r}
nonNative<-c("euonal","allipe_1","allipe_2","allipe_total")
community_ew_native<-community_ew[,!(names(community_ew) %in% nonNative)] 
```
#calculate total stem count and shannon diversity
```{r}
nativeShannon<-diversity(community_ew_native)
totalNativeStem<-rowSums(community_ew_native)
```
######## Separate out by functional group ######

##### Merge with site info and worms ####
```{r}
communityByPlot<-cbind(allPlants_just2Sites2017[,c(1:4)],community_ew)
```
#melt data so that I can merge with functional group
```{r}
melted<-melt(communityByPlot,id.vars=c("plotID","site","treatment","year"))
head(melted)
```
#remove funky double spaces in funcitonal groups look up table
```{r}
fxnLUT$functionalGroup<-gsub(" ", "", fxnLUT$functionalGroup, fixed = TRUE)
```
#merge melted data frame with functional gorup
```{r}
melted_fxn<-merge(melted,fxnLUT,by.x="variable",by.y="code",all.x=TRUE)
nrow(melted)==nrow(melted_fxn) #double check
```
#Reclassify tree and shrubs as woody
```{r}
melted_fxn$functionalGroup[melted_fxn$functionalGroup=="Tree"|melted_fxn$functionalGroup=="Shrub"]<-"Woody"
melted_fxn$functionalGroup[melted_fxn$functionalGroup=="Non-native"]<-"NonNative"
```
#drop allipe 1 and 2 and just keep total allipe
```{r}
melted_fxn<-melted_fxn[melted_fxn$variable!="allipe_1"&melted_fxn$variable!="allipe_2",]
unique(melted_fxn$variable) #double check
```
#summamrize by functional group
```{r}
fxnlGroupStemCounts<-melted_fxn %>% group_by(plotID,functionalGroup) %>% summarize(stemCount=sum(value))
```
#cast as columns
```{r}
fxnlGroupCasted<-dcast(fxnlGroupStemCounts,plotID~functionalGroup,value.var="stemCount",fun.aggregate = sum)
```
###### Put all data together ######
# Bind diversity with community data 
```{r}
bindNativeMetrics<-cbind(communityByPlot,nativeShannon,totalNativeStem)
```
#merge functional groups with bindNativeMetrics
```{r}
mergeFxnl<-merge(bindNativeMetrics,fxnlGroupCasted,by="plotID")
```
#merge with earthworm data
```{r}
allData<-merge(mergeFxnl,wormData,by="plotID")
nrow(allData) == 18
write.csv(allData, ".\\output\\garlicMustardEarthworm_prunseCot_to_fraxam_9April2018.csv",row.names=F)
```
############### Permanova ################
#get rid of nonNative
```{r}
allData_native<-allData[,!(names(allData) %in% nonNative)] 
```
#get bray distances using abundances of native plants (subset df by these columns)
```{r}
brayDist<-vegdist(allData_native[,c(5:35)]) #*will change if community data changes
myPerm<-adonis(brayDist ~treatment*mass, data = allData_native,strata=allData_native$site)
myPerm
```
#note that p values will change each time based on seed (random generator for permutaitons). I am not sure what seed I used in manuscript!
############### NMDS Plots ###########################
```{r}
meta<-metaMDS(allData_native[,c(5:35)]) 
```
#set parameters for NMDS
```{r}
treatment<- factor(allData$treatment)
col.gr <- c("red", "gray40", "black") #red corresponds to invaded, gray to pulled, black to unvivdaded
size<-log(allData$mass+1)*1.5 #for setting size of points proportional to worm biomass
```
#output
```{r}
svg(filename="nmds_withFrax.svg", 
    width=6.5, 
    height=6.5)
plot(meta, type = "n",cex.axis=1.8, cex.lab=1.8)
points(meta, display = "sites", cex = size, pch=19, col = col.gr[treatment])
```
#text(meta, display = "spec", cex=0.7)
```{r}
ordiellipse(meta, allData$site, conf = 0.95, label = TRUE,cex=1.8)
dev.off()
```


